---
author: "Hu Die"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction
There are four functions in this package accorded with the research based on the **Yang J, Lee SH, Goddard ME, Visscher PM (2011) GCTA: a tool for genome-wide complex trait analy- sis. Am J Hum Genet 88: 76–82. doi: 10.1016/j.ajhg.2010.11.011 PMID: 21167468**. These functions aim to simulate the family file, relationship file, genotype file and environment relationship matrix file. So that the outputs can be analyzed in the Variance Component Analysis in __GCTA__ from the reference paper.

## Functions 

#### fam_genera()
This function is to simulate the family file which is a data.frame output containing five columns: Family_ID, Individual_ID, Paternal_ID, Maternal_ID and Sex. With the input __n__, the number of the families, the output would generate the five kinds of information of __4n__ individuals for every family is set to consist of two parents and two children.

```{r,eval=FALSE}
fam_genera <- function(n){

  fid <- rep (1:n ,4 )
  iid <- c(rep(1,n),rep(2,n),rep(3,n),rep(4,n))
  iid <- paste0(fid,paste0("000",iid))
  pid <- rep(1,2*n)
  mid <- rep(2,2*n)
  pid <- c(rep(0,2*n),paste0(fid[1:(2*n)],paste0("000",pid)))
  mid <- c(rep(0,2*n),paste0(fid[1:(2*n)],paste0("000",mid)))
  sex <- c(rep(1,n),rep(2,n),sample(1:2,2*n,replace=TRUE))
  fam <- data.frame(fid,iid,pid,mid,sex)
  fam <- fam[order(fam$fid),]
  return(fam)
}
```

Set __n=50__, and __head(result)__:

```{r,eval=TRUE}
library(SC19036)
n <- 50
result <- fam_genera(n)
head(result)
```

#### fam_relation()
This function is to generate the family relationship based on the file by __fam_genera()__. With every family containing two parents and two children, there would be six pairs of relationships: "couple","parentOffspring","parentOffspring","parentOffspring", "parentOffspring" , "fullSib" standing for the relationship between the parents, between the parent and a child between and the children. And the output is a data.frame with four columns: Family_ID, Individual_ID1, Individual_ID2, Relationship. Then this out put could be used for generating the __environment relationship matrix__.

```{r,eval=FALSE}
fam_relation <- function (fam){
  n <- nrow(fam)
  fam[,1] <- as.numeric(as.character(fam[,1]))
  fam[,2] <- as.numeric(as.character(fam[,2]))
  fid <- rep(unique(fam[,1]),each=6)[1:(6*n/4)]
  iid1 <- iid2 <- relationship <- NULL
  for( i in 1:length(unique(fid))){
    
    a <- unique(fid)[i]
    l <- unique(fam[which(fam[,1]==a),2])
    iid1[(i-1)*6+1] <- l[1]
    iid2[(i-1)*6+1] <- l[2]
    iid1[(i-1)*6+2] <- l[1]
    iid2[(i-1)*6+2] <- l[3]
    iid1[(i-1)*6+3] <- l[1]
    iid2[(i-1)*6+3] <- l[4]
    iid1[(i-1)*6+4] <- l[2]
    iid2[(i-1)*6+4] <- l[3]
    iid1[(i-1)*6+5] <- l[2]
    iid2[(i-1)*6+5] <- l[4]
    iid1[(i-1)*6+6] <- l[3]
    iid2[(i-1)*6+6] <- l[4]
  }
  relationship <- rep(c("couple","parentOffspring","parentOffspring",
                        "parentOffspring", "parentOffspring" , "fullSib" ),length(unique(fid)))
  re <- data.frame(fid,iid1,iid2,relationship)
  
  return(re)
  
}
```

Use the __result__ generated by __fam_genera()__ to be the input here, and show the __head(relation)__.

```{r,eval=TRUE}
library(SC19036)
relation <- fam_relation(result)
head(relation)
```

#### geno_genera()
This function is to simulate the genotypes data. With the input family file and the number of SNPs, it would create the genotypes file. At each SNP, the minor allel frequency is sampled from the __runif()__ with __minimum = 0.1__. And the genotypes of every individual is regulated by the Mendel's Law of Genetics. The out put is a list here. the first element is the bed file consisting the original fam file, the phenotyps (by default set -9 as not known) and the genotypes data. Here each column of the genotypes is simplified as three classes: 0, 1, 2 standing for the number of the minor allels in the individual's genotype at this SNP. After changing one column of SNP here into two columns demonstrating whether the indivial owns the base at each SNP, then could the transcribed file be read as __.ped__ file in the tool __PLINK__. 

```{r,eval=FALSE}
geno_genera <- function (text , m){
  n <- length(unique(text[,1]))
  text[,1] <- as.numeric(as.character(text[,1]))
  text[,2] <- as.numeric(as.character(text[,2]))
  text[,3] <- as.numeric(as.character(text[,3]))
  text[,4] <- as.numeric(as.character(text[,4]))
  text[,5] <- as.numeric(as.character(text[,5]))
  outbed <- integer()
  pm <- runif(m,0.1,1)
  for(j in 1:n){
    no.fam <- unique(text[,1])[j]
    l.fam <- text[which(text[,1]==no.fam),2]
    iid <- l.fam[order(l.fam)]
    fid <- rep(no.fam,length(l.fam))
    bed <- cbind(fid, iid)
    for(i in 1:m){
      p <- pm[i]
      ## p : the MAF of one SNP
      
      AA <- (1-p)^2
      Aa <- 2*p*(1-p)
      aa <- p^2
      prob <- c(AA^2, AA*Aa,AA*aa,Aa*AA,Aa*Aa,Aa*aa,aa*aa)
      
      pairs <- 1:7
      parent <- sample (pairs, 1,prob=prob)
      if(parent==1) father <- mother <- kid1 <- kid2 <- 0
      if(parent==2) {
        father <- 0
        mother <- 1
        kid1 <- sample(0:1,1)
        kid2 <- sample(0:1,1)
      }
      if(parent==3){
        father <- 0
        mother <- 2
        kid1 <- kid2 <-1
      }
      if(parent==4){
        father <- 1
        mother <- 0
        kid1 <- sample(0:1,1)
        kid2 <- sample(0:1,1)
      }
      if(parent==5){
        father <- mother <- 1
        kid1 <- sample(0:2, 1, prob=c(0.25,0.5,0.25))
        kid2 <- sample(0:2, 1, prob=c(0.25,0.5,0.25))
      }
      if(parent==6){
        father <- 1
        mother <- 2
        kid1 <- sample (1:2 , 1)
        kid2 <- sample (1:2 , 1)
      }
      if(parent==7) father <- mother <- kid1 <- kid2 <-2
      snp <- c(father, mother ,kid1,kid2)
      bed <- cbind(bed,snp)
    }
    outbed <- rbind(outbed,bed)
  }
  colnames(outbed) <- c("fid","iid",paste0("snp",1:m))
  phenotype <- rep(-9,nrow(outbed))
  outbed0 <- data.frame(text[,1:5],phenotype,outbed[,3:ncol(outbed)])
  l <- list()
  l$bed <-outbed0
  l$maf <- pm
  return(l)
}

```

Here, use the __result__ by __fam_genera()__ and set __m=100__, then show the __head(geno)__.

```{r,eval=TRUE}
library(SC19036)
m <- 100
geno <- geno_genera(result,m)
head(geno[[1]],5)[,1:12]
head(geno[[2]],10)
```

#### ERM_genera()
This file is based on the idea put forward by **Xia, C. et al. Pedigree- and SNP-associated genetics and recent environment are the major contributors to anthropometric and cardiometabolic trait variation. PLoS. Genet. 12, e1005804 (2016)**. The idea is to make the environment relationship matrix from three different aspects: family aspect, an impact on both of the parent and the child; couple aspect, the impact on both of the couples, here the parents; fullsibling aspect, the impact on both of the fullsiblings. And for each environment aspect, we can generate the environment relationship matrix. With __n__ individuals all over, then we can get the matrix with n plus n. And the diagonal entry is 1 while the non-diagonal entry is 1 only if the row individual and the column individual is some relaship according to the environment relationship matrix. For example, the couple ERM would be 1 in the non-diagonal entry where the relathionship between the two individuals is couple relationship. 

And the output is a list consisting of six files. For each impact, there exist two files: one is the environment relationship matrix, the other one is the id information. And the matrix file could be saved and renamed as __.grm.gz__ to be used in the **GCTA** tool while the id file could be saved and renamed ad __.grm.id__ to be used in the **GCTA** tool. With that, the Variance Component analysis according to the model in **Xia(2016)** can be analyzed.

```{r,eval=FALSE}
ERM_genera <- function(text){
  fam_file <- text
  fam_file <- fam_file[, 1:2]
  names(fam_file) <- c("FID", "IID")
  fam_file$order <- 1:nrow(fam_file)
  rel <- re[, c( 2,3, ncol(re))]
  names(rel) <- c("IID_1", "IID_2", "relationship")
  rel2 <- rel[,c(2,1,3)]
  names(rel2) <- c("IID_1", "IID_2", "relationship")
  rel <- rbind(rel, rel2)
  
  rel <- merge(rel, fam_file[,c("IID", "order")], by.x="IID_1", by.y="IID")
  rel <- merge(rel, fam_file[,c("IID", "order")], by.x="IID_2", by.y="IID")
  rel <- rel[, c("IID_1", "IID_2", "relationship", "order.x", "order.y")]
  rel <- rel[order(rel$order.x, rel$order.y), ]
  sum(rel$order.x >= rel$order.y)
  rel <- rel[rel$order.x > rel$order.y, ]
  
  n <- nrow(fam_file)
  ### environment family
  rel$coef <- 0 
  rel[rel$relationship == "parentOffspring", "coef"] <- 1 ## hudie set  ##original 0.5
  rel[rel$relationship == "couple", "coef"] <- 1 ## hudie set  ##original 0.5
  rel[rel$relationship == "fullSib", "coef"] <- 1 ## hudie set  ##original 0.5
  x <- 1:n
  y <-1:n
  xy1 <- xy2 <- numeric()
  for(i in 1:(n-1)){
    for(j in (i+1):n) {
      xy1 <- c(xy1,x[i])
      xy2 <- c(xy2,y[j])
    }
  }
  xy <- cbind(xy1,xy2)
  for( i in 1:length(unique(rel$order.x))){
    a <- unique(rel$order.x)[i]
    b <- rel[which(rel$order.x==a),"order.y"]
    for(j in 1:length(b)){
      xy <- xy[-which(xy[,2]==a & xy[,1]==b[j]),]
    }
  }
  
  order.x <- c(rel$order.x,xy[,2])
  order.y <- c(rel$order.y,xy[,1])
  relationship <- c(rel$coef,rep(0,length(order.x)-length(rel$coef)))
  order.x <- c(order.x,1:n)##add the diagnal entry
  order.y <- c(order.y,1:n)
  
  relationship <- c(relationship,rep(1,n))
  snp <- rep(500,length(order.x))
  matrix_family <- data.frame(order.x, order.y, snp,relationship)
  matrix_family_id <- data.frame(order.x, order.y, snp,relationship)
  
  ###environment couple
  rel$coef <- 0 
  rel[rel$relationship == "couple", "coef"] <- 1 ##hudie-set 
  x <- 1:n
  y <- 1:n
  xy1 <- xy2 <- numeric()
  for(i in 1:(n-1)){
    for(j in (i+1):n) {
      xy1=c(xy1,x[i])
      xy2=c(xy2,y[j])
    }
  }
  xy <- cbind(xy1,xy2)
  for( i in 1:length(unique(rel$order.x))){
    a <- unique(rel$order.x)[i]
    b <- rel[which(rel$order.x==a),"order.y"]
    for(j in 1:length(b)){
      xy <- xy[-which(xy[,2]==a & xy[,1]==b[j]),]
    }
  }
  
  order.x <- c(rel$order.x,xy[,2])
  order.y <- c(rel$order.y,xy[,2])
  relationship <- c(rel$coef,rep(0,length(order.x)-length(rel$coef)))
  order.x <- c(order.x,1:n)##add the diagnal entry
  order.y <- c(order.y,1:n)
  
  relationship <- c(relationship,rep(1,n))
  snp <- rep(500,length(order.x))
  matrix_couple <- data.frame(order.x, order.y, snp,relationship)
  matrix_couple_id <- data.frame(order.x, order.y, snp,relationship)
  
  
  ##environment full-sibling
  rel$coef <- 0 
  rel[rel$relationship == "fullSib", "coef"] <- 1## hudie set  ##original 0.5
  x <- 1:n
  y <- 1:n
  xy1 <- xy2 <- numeric()
  for(i in 1:(n-1)){
    for(j in (i+1):n) {
      xy1 <- c(xy1,x[i])
      xy2 <- c(xy2,y[j])
    }
  }
  xy <- cbind(xy1,xy2)
  for( i in 1:length(unique(rel$order.x))){
    a <- unique(rel$order.x)[i]
    b <- rel[which(rel$order.x==a),"order.y"]
    for(j in 1:length(b)){
      xy <- xy[-which(xy[,2]==a & xy[,1]==b[j]),]
    }
  }
  
  order.x <- c(rel$order.x,xy[,2])
  order.y <- c(rel$order.y,xy[,2])
  relationship <- c(rel$coef,rep(0,length(order.x)-length(rel$coef)))
  order.x <- c(order.x,1:n)##add the diagnal entry
  order.y <- c(order.y,1:n)
  
  relationship <- c(relationship,rep(1,n))
  snp <- rep(500,length(order.x))
  matrix_fullsilb <- data.frame(order.x, order.y, snp,relationship)
  matrix_fullsilb_id <- data.frame(order.x, order.y, snp,relationship)
  
  l <- list(matrix_family,matrix_family_id,matrix_couple,matrix_couple_id,matrix_fullsilb,matrix_fullsilb_id)
  return(l)
  }

```














